Thesis outline

### 2017-06-15 KellyDouglass/Thesis/DirectMethodPhysics.tex
Edits to Abstract, Intro, Appendix

UV photons from young stars in an \ion{H}{2} region keep the interstellar gas 
partially ionized.  In an \ion{H}{2} region, a photon with an energy 
$E_\gamma \gtrsim 13.6\text{ eV}$ ionizes a hydrogen atom, producing an H$^+$ 
ion and an electron with some kinetic energy 
$K_{e^-} = E_\gamma - 13.6\text{ eV}$.  As the free electron moves around in the 
gas, it loses some of its kinetic energy as it collisionally excites other ions 
in the gas.  Because we are in a Str\"{o}mgren sphere (an ionized region or an 
\ion{H}{2} region), the cross-section for this electron scattering is much 
larger than the photoionization cross section, so the free electrons will 
thermalize quickly \citep{DeRobertis87}.  Eventually, the free electron 
recombines with an H$^+$ ion, forming atomic hydrogen.  The ions that are 
collisionally excited by the electron before its recombination eventually 
de-excite via forbidden transitions, emitting radiation that escapes the 
\ion{H}{2} region.

@@

The relative excitation rates of the two lines shown in Fig. 
\ref{fig:transitions_P1} are proportional to their statistical weights; thus, the 
ratio of the line intensities is a constant in the low-density limit 
\citep{Osterbrock89}.  In the high-density regime, this ratio is best accurately 
described by a Boltzmann population ratio.  There is a critical density for the 
energy levels which describes the turning point between these two extremes.  A 
graphical representation of this ratio as a function of density can be seen in 
Fig. \ref{fig:SII_den}.

>>>>

The Direct $T_e$ method is used to calculate the gas-phase chemical abundances 
(oxygen, nitrogen, etc. relative to hydrogen) in \ion{H}{2} regions.  It differs 
from other commonly-used methods by estimating the gas temperature and electron 
density of the region directly from the galaxy's spectrum.  UV photons from 
young stars in an \ion{H}{2} region keep the interstellar gas partially ionized.  
In an \ion{H}{2} region, a photon with an energy 
$E_\gamma \gtrsim 13.6\text{ eV}$ ionizes a hydrogen atom, producing an H$^+$ 
ion and an electron with some kinetic energy 
$K_{e^-} = E_\gamma - 13.6\text{ eV}$.  As the free electron moves around in the 
gas, it loses some of its kinetic energy as it collisionally excites other ions 
in the gas.  Because we are in a Str\"{o}mgren sphere (an ionized region or an 
\ion{H}{2} region), the cross-section for this electron scattering is much 
larger than the photoionization cross section, so the free electrons will 
thermalize quickly \citep{DeRobertis87}.  Eventually, the free electron 
recombines with an H$^+$ ion, forming atomic hydrogen.  The ions that are 
collisionally excited by the electron before its recombination eventually 
de-excite via forbidden transitions, emitting radiation that escapes the 
\ion{H}{2} region.

@@

The relative excitation rates of the two lines shown in Fig. 
\ref{fig:transitions_P1} are proportional to their statistical weights; thus, the 
ratio of the line intensities is a constant in the low-density limit 
\citep{Osterbrock89}.  In the high-density regime, this ratio is best accurately 
described by a Boltzmann population ratio.  There is a critical density for the 
energy levels which describes the turning point between these two extremes.  A 
graphical representation of this ratio as a function of density can be seen in 
Fig. \ref{fig:SII_den}.  \ion{H}{2} regions typically have densities well within 
the low-density regime, so we set the electron density for all our galaxies at 
100 cm$^{-3}$.

### 2017-06-15 KellyDouglass/Thesis/Thesis_sections/Introduction.tex
Edits to Abstract, Intro, Appendix

\begin{figure}
    \includegraphics[width=0.65\textwidth]{Images/Intro/sdss}
    \caption[SDSS galaxy map]{A slice through the SDSS galaxy distribution; each 
    point on this plot is a galaxy.  Large sky surveys such as SDSS have 
    revealed a non-uniform distribution of galaxies throughout the universe.  
    Dubbed the cosmic web, galaxies clump together in clusters and along 
    filaments, leaving giant gaps between them (similar to a sponge).}
    \label{fig:SDSS_map}
\end{figure}

\begin{figure}
    \includegraphics[width=0.9\textwidth]{Images/Intro/Millenium_simulation}
    \caption[Dark matter simulation]{Snapshots of a portion of the Millenium 
    Simulation showing the dark matter distribution at different points in time 
    from 0.21 Gyr after the Big Bang to the present.  Small perturbations in the 
    initial distribution of dark matter are amplified as the universe expands.  
    Gravity causes the slightly overdense regions to contract while 
    simultaneously causing the underdense regions to expand.}
    \label{fig:DMsim}
\end{figure}

Large sky surveys \citep[such as the Sloan Digital Sky Survey --- 
SDSS;][]{York00} have revealed a non-uniform distribution of galaxies throughout 
the universe.  Taking on a shape similar to that of a sponge or a 
three-dimensional cosmic web \citep{Bond96}, galaxies clump together in clusters 
and along filaments, leaving void regions between these large-scale structures.  
Evidence of this distribution can be seen in Fig. \ref{fig:SDSS_map} for 
galaxies in the SDSS Data Release 7 \citep[SDSS DR7;][]{Abazajian09}.  Based on 
these observations, dark matter simulations have been constructed which 
successfully reproduce the same large-scale structure.  If we start with a 
mostly uniform distribution of dark matter at the Big Bang, Fig. \ref{fig:DMsim} 
outlines the evolution of that dark matter up through the present.  These 
snapshots from the Millenium Simulation project \citep{Springel05} show that 
small perturbations in the initial distribution of dark matter are amplified as 
the universe expands.  Gravity causes the slightly overdense regions to collapse 
while simultaneously causing the underdense regions to expand.  Due to dark 
matter's strong gravitational interaction, we believe that the baryonic mattter 
(matter which interacts electromagnetically) will trace the dark matter 
distribution, resulting in the galaxy distribution we observe today.

@@

Existing in the cosmological voids, void galaxies are thought to demonstrate the 
fundamental characteristics of galactic evolution.  Cosmic voids are an 
important environment for studying galaxy formation \citep[see][for a review]
{vandeWeygaert11}.  Gravitational clustering proceeds as if in a very low 
density universe, where the amassment of gravitationally bound dark matter halos 
ends relatively early and there is little subsequent interaction between the 
galaxies due to the lower density and the faster local Hubble expansion.  
Therefore, $\Lambda$CDM cosmology predicts that galaxies formed in voids should 
have lower mass and be retarded in their star formation when compared to those 
in denser regions \cite[e.g.,][]{Gottlober03,Goldberg05,Cen11}.  
\cite{Goldberg04} show that a void region that is only 10\% of the mean density 
of the universe with $\Omega_{matter} = 0.3$, $h = 0.7$ dynamically evolves as 
if $\Omega_{matter} = 0.02$, $\Omega_\Lambda = 0.48$, and $h = 0.84$.  
Additionally, hydrodynamical cosmological simulations by \cite{Cen11} show that 
void galaxies continue to form stars because the gas in voids remains below the 
critical entropy threshold.  Void galaxies evolve in a relatively pristine 
environment where interactions are rare and star formation proceeds up to the 
present epoch because void galaxies are able to retain their gas.  This 
contrasts the denser environments, where the chemical composition and evolution 
of galaxies are drastically altered due to mergers and tidal and/or ram-pressure 
stripping.

% What do we already know about the environmental influence?
Observational studies of void galaxies have included investigation of 
photometric properties such as luminosity \citep{Hoyle05,Croton05,Moorman15}, 
color and morphological type \citep{Grogin00,Rojas04,Patiri06,Park07,
vonBendaBeckmann08,Hoyle12}, star formation rates estimated from optical 
spectroscopy and UV photometry \citep{Rojas05,Moorman15,Beygu16}, and gas 
content \citep{Kreckel12,Moorman16,Jones16}.  When compared to galaxies in 
denser regions, void galaxies are fainter, of late morphological types, bluer, 
forming stars at higher rates per unit stellar mass, and more gas rich.

@@

In particular, we study the effects of the large-scale environment on the 
formation and evolution of dwarf galaxies.  Fig. \ref{fig:Mr_SDSS} shows the 
distribution of galaxies by their absolute magnitudes detected in SDSS DR7.  
Defined to be galaxies with absolute magnitudes fainter than -17 ($M_r > -17$), 
the low stellar masses of these systems correspond to more shallow gravitational 
potential wells.  This causes any environmental effects to have a more 
significant influence on dwarf galaxies than on galaxies with larger stellar 
masses.  In addition, studying void dwarf galaxies will complement previous 
studies of dwarfs in groups and clusters because the assembly histories of 
low-mass galaxies are predicted to be very different \citep[e.g.,][]{Gao07,
Lackner12}.  Observations to date also show that the properties of dwarf 
galaxies vary dramatically with the environment \citep[e.g.,][]{Ann08,Geha12}.

@@

The first stars that formed are thought to have been composed of only hydrogen 
(and possibly some trace amount of helium).  As stars burn, nucleosythesis 
creates heavier elements at the core.  When the stars die and expel their 
content, the heavier elements are distributed between the surrounding 
interstellar gas and circumgalactic medium.  When new stars form from this gas, 
they include both hydrogen and some of the heavier elements.  By measuring the 
fraction of elements heavier than helium (``metals'') relative to the hydrogen 
content (the metallicity), we measure the integrated star formation history of 
the galaxy.

\begin{figure}
    \includegraphics[width=0.9\textwidth]{Images/Intro/Kirchoff}
    \caption[Kirchoff's laws of spectroscopy]{Three different viewpoints of the 
    light emitted from a star and/or a cool cloud of gas, representing 
    Kirchoff's laws of spectroscopy.  A star emits light across all wavelengths, 
    so the resulting spectrum is known as a continuous spectrum.  Star's light 
    that has passed through a cloud of cool gas before being observed is an 
    absorption spectrum, where the elements in the cloud have absorbed some of 
    the light at specific wavelengths (corresponding to the particular elements 
    present in the gas).  If observed off the line-of-sight of the star, the 
    light emitted from the cloud of cooler gas results in an emission spectrum, 
    where the gas is re-radiating the light it absorbed from the star.}
    \label{fig:Kirchoff}
\end{figure}

We use a galaxy's spectrum to measure various properties of the stellar 
population and interstellar medium.  A galaxy's spectrum is a superposition of 
light from the stars and emission lines from the gas.  Depending on our 
viewpoint of the galaxy's light, we will see a different type of spectrum.  As 
shown in Fig. \ref{fig:Kirchoff}, if we separate a star's light by its 
wavelength, we will observe a continuous spectrum, since it is a blackbody.  If 
instead we observe the starlight after it has passed through a cloud of cooler 
gas, we will measure an absorption line spectrum; the elements in the gas cloud 
have absorbed some of the star's light at specific wavelengths corresponding to 
the particular ions present.  Finally, if we only observe the light emitted from 
the cool gas cloud, we will measure an emission line spectrum.  The method we 
employ to estimate the metallicity of a galaxy requires particular emission 
lines.  We study the gas-phase chemical abundances of \ion{H}{2} regions in a 
galaxy, where the light is coming from low-density gas partially ionized by UV 
photons emitted from young, hot stars.  Therefore, we are measuring the 
abundances of the elements produced in previous generations of stars.

\begin{figure}
    \includegraphics[width=0.9\textwidth]{Images/Intro/spectrum}
    \caption[Sample (void) dwarf galaxy spectrum]{Example SDSS DR7 spectrum --- 
    this is a void dwarf galaxy whose gas-phase chemical abundances are slightly 
    above average.  The [\ion{O}{3}] $\lambda$4363 auroral line is circled in 
    red, while the [\ion{O}{3}] $\lambda \lambda$4959,5007 doublet is circled in 
    yellow.}
    \label{fig:spectrum}
\end{figure}

The ion transitions observed in the optical part of the spectrum (where SDSS DR7 
operates) consist of forbidden transitions.  As explained in detail in Section 
\ref{sec:physics}, forbidden transitions have much longer lifetimes than the 
allowed electron transitions.  These transitions depend strongly on the 
temperature and density of the electron gas, because the long lifetimes allow 
the atoms to easily collisionally de-excite.  By taking the ratio of the 
[\ion{O}{3}] $\lambda$4363 auroral line to the [\ion{O}{3}] 
$\lambda \lambda$4959,5007 doublet, we can get an estimate of the temperature of 
the gas.  These lines are highlighted on the example SDSS DR7 spectrum shown in 
Fig. \ref{fig:spectrum}.

Oxygen is used as a proxy to measure the metallicity of an \ion{H}{2} region, 
because oxygen is the most abundant element in the universe after hydrogen and 
helium.  It also has strong emission lines in the visible wavelength range, so 
it is relatively easy to observe.  Oxygen emission is the most effective cooling 
source in an \ion{H}{2} region, so lower metallicities correspond to higher 
temperatures.  There are three main factors which determine the strength of an 
emission line: the temperature, density, and number of ions in the gas.  It is 
possible to calculate the number ratios for any of the heavy elements with 
respect to hydrogen once the electron temperature and density of the 
interstellar gas are estimated.

% Outline of Thesis
% Pose the problem - how does the large-scale environment affect galaxy evolution?
We want to understand how dark matter affects the evolution of a galaxy and what 
influence it has on a galaxy's star formation.  To do this, we compare the 
properties of galaxies living in voids with galaxies in denser regions in an 
effort to understand how the properties of the gas and the history of star 
formation in a galaxy depend on the environment.  We then try to infer what 
these results tell us about the dark matter structure and history of the galaxy.  
In Chapter \ref{ch:Paper1}, we calculate and compare the metallicity of 135 
star-forming dwarf galaxies as a function of their large-scale environment.  We 
continue studying this galaxy sample in Chapter \ref{ch:Paper2} by estimating 
their ratios of nitrogen to oxygen and looking at how that depends on the 
large-scale environment.  To expand our dwarf galaxy sample, we derive an 
approximation for the amount of singly-ionized oxygen and repeat our analysis of 
the large-scale environmental influence on the gas-phase chemical abundances of 
1920 star-forming dwarf galaxies in Chapter \ref{ch:Paper3}.  In Chapter 
\ref{ch:smallScaleEnvironment}, we investigate how the dwarf galaxies' 
properties depend on the galaxies' distances to the nearest neighbors and 
groups.  Finally, in an effort to better understand how galaxies evolve from 
blue, star-forming systems to ``red and dead,'' we discover a way to classify 
galaxies into their location on the color-magnitude diagram in Chapter 
\ref{ch:GV}.

>>>>

\begin{figure}
    \includegraphics[width=0.65\textwidth]{Images/Intro/sdss}
    \caption[SDSS galaxy map]{A slice through the SDSS galaxy 
    distribution\footnote{\url{http://www.sdss.org/science/orangepie/}}; each 
    point on this plot is a galaxy.  Large sky surveys such as SDSS have 
    revealed a non-uniform distribution of galaxies throughout the universe.  
    Dubbed the cosmic web, galaxies clump together in clusters and along 
    filaments, leaving giant gaps between them (similar to a sponge).}
    \label{fig:SDSS_map}
\end{figure}

\begin{figure}
    \includegraphics[width=0.9\textwidth]{Images/Intro/Millenium_simulation}
    \caption[Dark matter simulation]{Snapshots of a portion of the Millenium 
    Simulation\footnote{\url{https://wwwmpa.mpa-garching.mpg.de/galform/virgo/millennium/}} 
    showing the dark matter distribution at different points in time from 0.21 
    Gyr after the Big Bang to the present.  Small perturbations in the initial 
    distribution of dark matter are amplified as the universe expands.  Gravity 
    causes the slightly overdense regions to contract while simultaneously 
    causing the underdense regions to expand.}
    \label{fig:DMsim}
\end{figure}

Large galaxy surveys such as the Sloan Digital Sky Survey \citep[SDSS;][]
{York00} have revealed a non-uniform distribution of galaxies throughout 
the universe.  Taking on a shape similar to that of a sponge-like topology 
\citep{Gott98} or a three-dimensional cosmic web \citep{Bond96}, galaxies clump 
together in clusters and along filaments, leaving void regions between these 
large-scale structures.  Evidence of this distribution can be seen in Fig. 
\ref{fig:SDSS_map} for galaxies in the SDSS Data Release 7 \citep[SDSS DR7;][]
{Abazajian09}.  Based on these observations, dark matter simulations have been 
constructed which successfully reproduce the same large-scale structure.  If we 
start with a mostly uniform distribution of dark matter at the Big Bang with 
quantum mechanical perturbations, Fig. \ref{fig:DMsim} outlines the evolution of 
that dark matter up through the present.  These snapshots from the Millenium 
Simulation project \citep{Springel05} show that small perturbations in the 
initial distribution of dark matter are amplified as the universe expands.  
Gravity causes the slightly overdense regions to collapse while simultaneously 
causing the underdense regions to expand.  Due to dark matter's gravitational 
attraction, we believe that the baryonic mattter (matter which interacts 
electromagnetically) will trace the dark matter distribution, resulting in the 
galaxy distribution we observe today.

@@

Existing in the cosmological voids, void galaxies are thought to demonstrate the 
fundamental characteristics of galactic evolution.  Cosmic voids are an 
important environment for studying galaxy formation \citep[see][for a review]
{vandeWeygaert11}.  Gravitational clustering proceeds as if in a very low 
density universe, where the amassment of gravitationally bound dark matter halos 
ends relatively early and there is little subsequent interaction between the 
galaxies due to the lower density and the faster local Hubble expansion.  
Therefore, $\Lambda$CDM cosmology predicts that galaxies formed in voids should 
have lower mass and be retarded in their star formation when compared to those 
in denser regions \cite[e.g.,][]{Gottlober03,Goldberg05,Cen11}.  
\cite{Goldberg04} show that a void region that is only 10\% of the mean density 
of the universe with $\Omega_{matter} = 0.3$, $h = 0.7$ dynamically evolves as 
if $\Omega_{matter} = 0.02$, $\Omega_\Lambda = 0.48$, and $h = 0.84$.  
Additionally, hydrodynamical cosmological simulations by \cite{Cen11} show that 
void galaxies may continue to form stars because the gas in voids remains below 
the critical entropy threshold (above which the cooling time exceeds the Hubble 
time).  Void galaxies evolve in a relatively pristine environment where 
interactions are rare and star formation proceeds up to the present epoch 
because void galaxies are able to retain their gas.  This contrasts with denser 
environments, where the chemical composition and evolution of galaxies are 
drastically altered due to mergers and tidal and/or ram-pressure stripping.

% What do we already know about the environmental influence?
Observational studies of void galaxies have included investigation of 
photometric properties such as luminosity \citep{Hoyle05,Croton05,Moorman15}, 
color and morphological type \citep{Grogin00,Rojas04,Patiri06,Park07,
vonBendaBeckmann08,Hoyle12}, star formation rates estimated from optical 
spectroscopy and UV photometry \citep{Rojas05,Moorman15,Beygu16}, and gas 
content \citep{Kreckel12,Moorman16,Jones16}.  When compared to galaxies in 
denser regions, void galaxies tend to be fainter, of late morphological types, 
bluer, forming stars at higher rates per unit stellar mass, and more gas rich.

@@

In particular, we study the effects of the large-scale environment on the 
formation and evolution of dwarf galaxies.  Fig. \ref{fig:Mr_SDSS} shows the 
distribution of galaxies by their absolute magnitudes detected in SDSS DR7.  
Defined to be galaxies with absolute magnitudes fainter than -17 ($M_r > -17$), 
the low stellar masses of dwarf galaxies correspond to more shallow 
gravitational potential wells.  This causes any environmental effects to have a 
more significant influence on dwarf galaxies than on galaxies with larger 
stellar masses.  In addition, studying void dwarf galaxies will complement 
previous studies of dwarfs in groups and clusters because the assembly histories 
of low-mass galaxies are predicted to be very different \citep[e.g.,][]{Gao07,
Lackner12}.  Observations to date also show that the properties of dwarf 
galaxies vary dramatically with the environment \citep[e.g.,][]{Ann08,Geha12}.

@@

The first stars that formed are thought to have been composed of hydrogen and 
helium (and trace amounts of other primordial elements).  As stars burn, 
nucleosythesis creates heavier elements at the core.  When the stars expel their 
contents in supernovae and stellar winds, the heavier elements are distributed 
between the surrounding interstellar gas and circumgalactic medium.  By 
measuring the ratio of elements heavier than helium (``metals'') relative to the 
hydrogen content, we measure the integrated star formation history of the 
galaxy.  This ``metallicity'' is commonly expressed as \OH, although it is 
sometimes given in units of the solar metallicity, $Z/Z_\odot$.

\begin{figure}
    \includegraphics[width=0.9\textwidth]{Images/Intro/Kirchoff}
    \caption[Kirchoff's laws of spectroscopy]{Three different viewpoints of the 
    light emitted from a star and/or a cool cloud of gas, representing 
    Kirchoff's laws of 
    spectroscopy\footnote{\url{http://courses.atlas.illinois.edu/spring2011/astr/astr210/LECTURES/Lect12.html}}.  
    A star emits light across all wavelengths, so the resulting spectrum is 
    known as a continuous spectrum.  Star's light that has passed through a 
    cloud of cool gas before being observed is an absorption spectrum, where the 
    elements in the cloud have absorbed some of the light at specific 
    wavelengths (corresponding to the particular elements present in the gas).  
    If observed off the line-of-sight of the star, the light emitted from the 
    cloud of cooler gas results in an emission spectrum, where the gas is 
    re-radiating the light it absorbed from the star.}
    \label{fig:Kirchoff}
\end{figure}

We use a galaxy's spectrum to measure various properties of the stellar 
population and interstellar medium.  A galaxy's spectrum is a superposition of 
light from the stars and emission lines from the interstellar gas.  The 
observer's viewpoint determines the type of spectrum measured.  As shown in Fig. 
\ref{fig:Kirchoff}, if we separate a star's light by its wavelength, we will 
observe a continuous spectrum, since its photosphere is close to a blackbody.  
If instead we observe the starlight after it has passed through a cloud of 
cooler gas (relative to the temperature of the star), we will measure an 
absorption line spectrum; the elements in the gas cloud have absorbed some of 
the star's light at specific wavelengths corresponding to the particular ions 
present.  Finally, if we only observe the light emitted from the cool gas cloud, 
we will measure an emission line spectrum.  The method we employ to estimate the 
metallicity of a galaxy requires the detection of particular emission lines.  We 
study the gas-phase chemical abundances of \ion{H}{2} regions in a galaxy, where 
the light is coming from low-density gas partially ionized by UV photons emitted 
from young, hot stars.  Therefore, we are measuring the abundances of the 
elements produced in previous generations of stars.

\begin{figure}
    \includegraphics[width=0.9\textwidth]{Images/Intro/spectrum}
    \caption[Sample (void) dwarf galaxy spectrum]{Example SDSS DR7 
    spectrum\footnote{\url{http://cas.sdss.org/dr7/en/get/specById.asp?ID=124634771904004096}} 
    --- this is a void dwarf galaxy whose gas-phase chemical abundances are 
    slightly above average.  The [\ion{O}{3}] $\lambda$4363 auroral line is 
    circled in red, while the [\ion{O}{3}] $\lambda \lambda$4959,5007 doublet is 
    circled in yellow.}
    \label{fig:spectrum}
\end{figure}

The ion transitions observed in the optical part of the spectrum (where SDSS DR7 
operates) consist of forbidden transitions.  As explained in detail in Section 
\ref{sec:physics}, forbidden transitions have much longer lifetimes than the 
allowed electric dipole electron transitions.  These transitions strongly depend 
on the temperature and density of the electron gas, because the long lifetimes 
allow the atoms to easily collisionally de-excite.  By taking the ratio of the 
[\ion{O}{3}] $\lambda$4363 auroral line to the [\ion{O}{3}] 
$\lambda \lambda$4959,5007 doublet, we can get an estimate of the temperature of 
the gas.  These lines are highlighted on the example SDSS DR7 spectrum shown in 
Fig. \ref{fig:spectrum}.

Oxygen is used as a proxy to measure the metallicity of an \ion{H}{2} region, 
because oxygen is the most abundant element in the universe after hydrogen and 
helium \citep{Emsley11}.  It also has strong emission lines in the visible 
wavelength range, so it is relatively easy to observe.  Oxygen emission is the 
most effective cooling source in an \ion{H}{2} region, so lower metallicities 
correspond to higher temperatures.  There are three main factors which determine 
the strength of an emission line: the gas temperature, the electron density, and 
the relative number of ions in the gas.  It is possible to calculate the number 
ratios for any of the heavy elements with respect to hydrogen once the electron 
temperature and density of the interstellar gas are estimated.

% Outline of Thesis
% Pose the problem - how does the large-scale environment affect galaxy evolution?
We want to understand how the cosmic environment affects the evolution of a 
galaxy and what influence it has on a galaxy's star formation.  To do this, we 
compare the properties of galaxies living in voids with galaxies in denser 
regions in an effort to understand how the properties of the gas and the history 
of star formation in a galaxy depend on the environment.  We then try to infer 
what these results tell us about the dark matter structure and history of the 
galaxy.  In Chapter \ref{ch:Paper1}, we calculate and compare the metallicity of 
135 star-forming dwarf galaxies as a function of their large-scale environment.  
We continue studying this galaxy sample in Chapter \ref{ch:Paper2} by estimating 
their ratios of nitrogen to oxygen and looking at how that depends on the 
large-scale environment.  To expand our dwarf galaxy sample, we derive an 
approximation for the amount of singly-ionized oxygen and repeat our analysis of 
the large-scale environmental influence on the gas-phase chemical abundances of 
1920 star-forming dwarf galaxies in Chapter \ref{ch:Paper3}.  In Chapter 
\ref{ch:smallScaleEnvironment}, we investigate how the dwarf galaxies' 
properties depend on the galaxies' distances to the nearest neighbors and 
groups.  Finally, in an effort to better understand how galaxies evolve from 
blue, star-forming systems to ``red and dead,'' we discover a way to classify 
galaxies into their location on the color-magnitude diagram in Chapter 
\ref{ch:GV}.  We present our conclusions and suggestions for future research in 
Chapter \ref{ch:conclusion}.

### 2017-06-15 KellyDouglass/Thesis/preamble.tex
Edits to Abstract, Intro, Appendix

% Papers 1-3
We investigate how the cosmic environment affects galaxy evolution in the 
Universe by studying gas-phase chemical abundances and other galaxy properties 
as a function of the local and cosmic density of galaxies.  Using spectroscopic 
observations from the Sloan Digital Sky Survey Data Release 7, we estimate the 
oxygen and nitrogen abundances of 993 star-forming void dwarf galaxies and 759 
star-forming dwarf galaxies in denser regions.  We use the Direct $T_e$ method 
for calculating the gas-phase chemical abundances in the dwarf galaxies because 
it is best suited for low metallicity, low mass galaxies.  A substitute for the 
[\ion{O}{2}] $\lambda$3727 doublet is developed, permitting oxygen abundance 
estimates of SDSS dwarf galaxies at all redshifts with the Direct $T_e$ method.  
We find that star-forming void dwarf galaxies have slightly higher oxygen 
abundances than star-forming dwarf galaxies in denser environments, but we find 
that void dwarf galaxies have slightly lower nitrogen abundances and lower N/O 
ratios than galaxies in denser regions.  
% Small-scale enivronment
At smaller scales, we find that only the presence of a neighboring galaxy within 
0.05 \hMpc or 0.05$r_{vir}$, or the presence of a group within 0.1 \hMpc, 
influences a dwarf galaxy's evolution.  Dwarf galaxies within 0.05 \hMpc or 
0.05$r_{vir}$ of another galaxy tend to be bluer, have higher sSFRs, and have 
higher oxygen and nitrogen abundances than average.  In contrast, galaxies 
within 0.1 \hMpc of the center of the closest group are redder, have lower 
oxygen abundances, and have higher N/O ratios than average.
% GV
We also investigate how a galaxy transitions through the color-magnitude 
diagram, evolving from a blue, star-forming spiral or irregular galaxy in the 
blue sequence to a red elliptical galaxy in the red cloud through the green 
valley.  We discover that combining a galaxy's color, color gradient, and 
inverse concentration index defines a galaxy's location on the color-magnitude 
diagram.  The results indicate that there is a higher fraction of dwarf galaxies 
in denser regions than void dwarf galaxies in the green valley.
% Conclusions
From these analyses, we surmise that void dwarf galaxies experience delayed star 
formation as predicted by the $\Lambda$CDM cosmology and a dependence of the 
large-scale environment on cosmic downsizing.  We present evidence that void 
dwarf galaxies have a higher ratio of dark matter halo mass to stellar mass when 
compared to dwarf galaxies in denser environments.

>>>>

% Papers 1-3
We investigate how the cosmic environment affects galaxy evolution in the 
Universe by studying gas-phase chemical abundances and other galaxy properties 
as a function of the large-scale environment and local density of galaxies.  
Using spectroscopic observations from the Sloan Digital Sky Survey Data Release 
7, we estimate the oxygen and nitrogen abundances of 993 star-forming void dwarf 
galaxies and 759 star-forming dwarf galaxies in denser regions.  We use the 
Direct $T_e$ method for calculating the gas-phase chemical abundances in the 
dwarf galaxies because it is best suited for low metallicity, low mass galaxies.  
A substitute for the [\ion{O}{2}] $\lambda$3727 doublet is developed, permitting 
oxygen abundance estimates of SDSS dwarf galaxies at all redshifts with the 
Direct $T_e$ method.  We find that star-forming void dwarf galaxies have 
slightly higher oxygen abundances than star-forming dwarf galaxies in denser 
environments, but we find that void dwarf galaxies have slightly lower nitrogen 
abundances and lower N/O ratios than galaxies in denser regions.  
% Small-scale enivronment
At smaller scales, we find that only the presence of a neighboring galaxy within 
0.05 \hMpc or 0.05$r_{vir}$, or the presence of a group within 0.1 \hMpc, 
influences a dwarf galaxy's evolution.  Dwarf galaxies within 0.05 \hMpc or 
0.05$r_{vir}$ of another galaxy tend to be bluer, have higher sSFRs, and have 
higher oxygen and nitrogen abundances than average.  In contrast, galaxies 
within 0.1 \hMpc of the center of the closest group are redder, have lower 
oxygen abundances, and have higher N/O ratios than average.
% GV
We also investigate how a galaxy transitions through the color-magnitude 
diagram, evolving from a blue, star-forming spiral or irregular galaxy in the 
blue sequence to a red elliptical galaxy in the red cloud through the green 
valley.  We discover that combining a galaxy's color, color gradient, and 
inverse concentration index defines a galaxy's location on the color-magnitude 
diagram.  The results indicate that there is a higher fraction of dwarf galaxies 
in denser regions than void dwarf galaxies in the green valley.
% Conclusions
From these analyses, we surmise that void dwarf galaxies experience delayed star 
formation as predicted by the $\Lambda$CDM cosmology.  We also conjecture that 
there is a dependence of the large-scale environment on cosmic downsizing.  We 
present evidence that void dwarf galaxies may have a higher ratio of dark matter 
halo mass to stellar mass when compared to dwarf galaxies in denser 
environments.

### 2017-06-15 KellyDouglass/Thesis/simulations.tex
Edits to Abstract, Intro, Appendix

\section{General dwarf galaxy formation and evolution}

@@

% ISM - inflows and outflows
The inflow and outflow of gas in galaxies are aspects of galaxy evolution that 
are not yet well understood.  This is especially important when studying dwarf 
galaxies, since their gravitational potential wells may be shallow enough to be 
strongly influenced by star formation and/or supernovae.  Simulations by 
\cite{Marcolini04} find that when ram pressure from the intergalactic medium 
(IGM) is comparable to the thermal pressure of the central interstellar medium 
(ISM), stripping and superwind influence each other and increase the gas removal 
rate.  They find that the amount of metal-rich ejecta is sensitive to the ram 
pressure.  
% supernova
Simulations by \cite{Power14} and \cite{Melioli15} show that supernovae contain 
sufficient energy to unbind the gas in a low mass halo.  \cite{Hu16} show that 
supernova-driven galactic outflows push metal-rich gas into the halo.  A small 
fraction of the gas eventually escapes the halo, while most falls back onto the 
disc.  This results in the halo metallicity being about 20\% higher than the 
disc metallicity.  Likewise, \cite{Muratov17} find that almost all metals 
produced in a Type II supernova are ejected from the galaxy.
% star formation
Unlike the supernovae, \cite{Melioli15} finds that the ISM material does not 
have a strong likelihood of being expelled from a galaxy due to star formation.  
In contrast, \cite{Muratov15} finds that a large portion of material is ejected 
in galactic outflows after a bust of star formation, which collects in the 
circumgalactic medium (CGM) and is sometimes recycled in later star formation 
episodes.  In galaxies with $M_h < 10^{12} M_\odot$, a fraction of material is 
lost to the IGM.

@@

\section{Environmental differences}

>>>>

\section{Dwarf galaxy formation and evolution}

@@

% ISM - inflows and outflows
The inflow and outflow of gas in galaxies are aspects of galaxy evolution that 
are not yet well understood.  This is especially important when studying dwarf 
galaxies, since their gravitational potential wells may be shallow enough to be 
strongly influenced by stellar winds and/or supernovae.  Simulations by 
\cite{Marcolini04} find that when ram pressure from the intergalactic medium 
(IGM) is comparable to the thermal pressure of the central interstellar medium 
(ISM), stripping and superwind influence each other and increase the gas removal 
rate.  They find that the amount of metal-rich ejecta is sensitive to the ram 
pressure.  
% supernova
Simulations by \cite{Power14} and \cite{Melioli15} show that supernovae contain 
sufficient energy to unbind the gas in a low mass halo.  \cite{Hu16} show that 
supernova-driven galactic outflows push metal-rich gas into the halo.  A small 
fraction of the gas eventually escapes the halo, while most falls back onto the 
disc.  This results in the halo metallicity being about 20\% higher than the 
disc metallicity.  Likewise, \cite{Muratov17} find that almost all metals 
produced in a Type II supernova are ejected from the galaxy.
% star formation
Unlike the supernovae, \cite{Melioli15} finds that the ISM material does not 
have a strong likelihood of being expelled from a galaxy due to star formation.  
In contrast, \cite{Muratov15} finds that a large portion of material is ejected 
in galactic outflows after a bust of star formation, which collects in the 
circumgalactic medium (CGM) and is sometimes recycled in later star formation 
episodes.  In galaxies with $M_h < 10^{12} M_\odot$, a fraction of material is 
lost to the IGM.

@@

\section[Environmental differences]{Environmental differences of galaxy formation and evolution}
